---
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Getting familiar with R programming

**This first chapter summarized all the basic information for understanding how R works.**

## **Basic understanding**

Before starting with Rstudio functioning, it is essential to briefly summarize the terminology and different elements used in programming.

### *Terminology*

-   **Command:** The command refers to a specific instruction or operation written in the R programming language that users enter into the console or the script editor. In other words, the command is the code that is written to perform various tasks.
-   **Function:** A function is a block of organized, reusable code designed to perform a specific task. Functions in R are used to encapsulate a set of operations or calculations that can be applied to input data, producing an output. They can be built-in functions provided by R or user-defined functions created by the programmer.
-   **Argument:** An argument is a value passed to a function to customize its behavior or provide input data. Functions in R can have zero or more arguments, which are specified within parentheses when the function is called. Arguments can be variables, constants, or expressions, and they influence the outcome of the function's execution.
-   **Script:** the script refers to a file containing a series of R commands and functions written in the R programming language. Scripts are used to organize and execute multiple commands sequentially to achieve specific tasks, such as data analysis, visualization, or modeling.

### *R studio elements*

-   **Console:** the console is the interactive environment where users can directly enter and execute R commands. It provides immediate feedback, displaying the output of commands, errors, warnings, and messages generated during code execution. It is important to remark that the data displayed in the console is not saved.
-   **Script Editor:** The script editor is where you write, edit, and save R scripts. Scripts contain a series of R commands and functions organized to perform specific tasks.
-   **Environment Pane:** this pane displays information about the current R environment, including data objects, functions, and variables that are currently loaded into memory.
-   **File Pane:** the file pane allows you to navigate your file system, view files, and manage your R projects. You can also create new files, folders, and scripts from this pane.
-   **Plots Pane:** when you create visualizations or plots in R, they are displayed in the plots pane. This pane allows you to view and interact with your plots.
-   **Packages Pane:** R packages are collections of functions, data sets, and documentation that extend the capabilities of R. The packages pane allows you to manage and install packages from CRAN (Comprehensive R Archive Network) and other sources.
-   **Help Pane:** this pane provides access to R documentation and help files. You can search for functions, packages, and topics to find relevant documentation and examples.
-   **Viewer Pane:** the viewer pane displays web-based content generated by R packages or Shiny applications.
-   **Git Pane:** if you are using version control with Git, the Git pane allows you to manage your Git repositories, commit changes, and view version history.
-   **Toolbar:** the toolbar provides quick access to commonly used actions and commands, such as saving files, running code, and debugging.

## **Basic operating**

### *Ins and outs*

In Rstudio there are two ways of running a command, in other words telling the system to do a task: we can press "ctrl" "Enter", or we can click on the "Run" icon in the tool bar. We can also cancel a command by pressing "ctrl" "c".

Rstudio will take everything that we write in the Script Editor or console as a command, and if we want to add a comment in our command we should write "\#" before, R will understand that what comes after is not a command. When we command something in the console, the results appear with a number in brackets, for example:

```{r}
1+1 #Here we are asking R to calculate the sum of 1 and 1
```

As can be seen in the example above, when we command "1+1" we get the result is "2". However, there is "\[1\]" before it. The number in brackets enumerates the values from the results, in this case, "2" is the first value from our results. This example also shows that the hashtag allows us to include comments in our script which will help to understand it better.

RStudio works best with vectorized data. A **vector** is a basic data structure that represents an ordered collection of elements of the same data type, and it can for example be a one-dimension set of numbers, such as the one bellow:

```{r}
1:6
```

In this example, we have used the operator ":" which gives as a result a group of numbers between 1 and 6. This group of number is what we call a vector.

If we would like to reuse this vector, then we would need to first save them. For this, Rstudio allows us to generate an **object**. In the context of RStudio, an object refers to a data structure that holds a value or values. We need to be careful when naming objects, if they start with a number or a special symbol Rstudio will not be able to run it. Also, we need to keep in mind that Rstudio is affected by the capital letters and "House" and "house" are two different things for R. Moreover, we can't forget that R overwrites the object names, we can always which names have been already assigned by running 1s()

Following up with the previous example we can create an object for the vector, so later on, we will be able to use it in other operations.

```{r}
a <- 1:6 #we generate the object named "a" and it will contain the result of the operation "1:6"
a
```

Now we will have an object named "a" in our environment which contains the vector (1, 2, 3, 4, 5, 6), and if we want to operate with it we just need to write the name of the object (a) in our command, such as:

```{r}
a+2
```

As it can be seen in the example, when we sum 2 to the vector named "a", R does **element-wise execution**, in other words, it sums 2 to every element in our vector, in this example: (1+2, 2+2, 3+2, 4+2, 5+2, 6+2). R will also element-wise execution when operating with two vectors, and if these do not have the same lengths, R will repeat the short one as long as the long one, this is known as **vector recycling**.

However, if we would like to do a traditional matrix operation, we should include "%" before and after our operation, for example:

```{r}
a%*%a
```

### *Functions*

A function is the task that we ask R to realize, it contains the operations or calculations that can be applied to input data. Some of examples of functions can be the ones bellow:

-   **Round()** it is used to round numerical values to a specified number of decimal places, for example:

```{r}
round(2,1)
```

-   **mean()** it is used to calculate the arithmetic mean (average) of a numeric vector or values, for example:

```{r}
mean(a)
```

-   **sum()** it is used to calculate the sum of numeric values in a vector or a data frame, for example:

```{r}
sum(a)
```

-   **sample()** it is used to generate random samples from a specified set of elements, for example:

```{r}
sample(a)
```

-   **c()** it is used to combine values into a single vector or list, for example:

```{r}
c(1, 5, 12, 20)
```

There are many functions that we can use in Rstudio, and all of them are composed by three main parts: a name, a body of the code and a set of arguments, for instance the one below:

```{r}
ana <- function(){
  b <- c(1, 2, 3, 4, 5, 6)
  b <- sample(b, size = 2)
  sum(b)
}
ana ()
```

In this example the name of the function is "ana", the body is everything that come in between the ""{}" and the arguments in this case are values that input for the functions, such as in the case of "sample()" the arguments are "b" (the data set that we are imputing) and "size =2" (how many values do we want that R gets out of this data set). Note that if we don't know which are the arguments for a function we can always use "args(the name of the function)2 in the console and it will give us a description how to imput the arguments. Now that we have name the function we can just write the name instead of rewriting the entire function body, as it can be seen in the example"ana()" gives the output of the entire function.

We also put an argument in the function, such as "ana(b2)", now the function "ana" will only work when we input what the argument "b2" is, for instance:

```{r}
ana <- function(b2){
  b3 <- sample(b2, size = 2)
  sum(b3)
}
#ana() gives us error because we have not defined what "b2" is
```

```{r}
ana <- function(b2){
  b3 <- sample(b2, size = 2)
  sum(b3)
}
ana (b2 = 1:4) #Here we define what "b2" is and Rstudio can run the function "ana()" for it
```

?apply #Note for myself: look it up \### *Packages*

We can build up our own functions, but can also use these that other people have already coded. For this, we have to install **packages** and load their **libraries**. A package is a collection of R functions, data sets, and compiled code in a well-defined format, while a library refers to a directory where installed packages are stored on your computer. When you install a package, its files are copied to a specific location on your system. This location is referred to as the library. So, depending on our aim we will install different packages and load their libraries. Note that we only need to install a package once, but we need to load its library every R session we want to use it.

For instance, "qplot" is a function used for making quick plots in R, but it is part of the package"ggplot2", so for being able to use it, we must install the package and load the library:

```{r}
#install.packages("ggplot2") to install the package (I don't want to run because I alredy have it)
library(ggplot2) #load the function
x <- c(1,1,2,2,2,3) #define the data
qplot(x, binwidth = 1) #the function and the arguments we want to use
```

Each package uses different functions, and it might be confusing to know how they operate and how to input the augments. R comes with a "Help" option, if we write in the console "?packagename", where packagename refers to the actual name of the package, we will will the information about how to use it. Note that for being able to get the help, the package need to be installed before.
